Paper on the carbonate chemistry at Ny-Ålesund

### 2020-04-20 paper.Rmd
?

d <- d_all %>%
  dplyr::filter(year < 2020) %>%
  dplyr::mutate(month=month(datetime, label = TRUE, abbr = TRUE)) # abbreviated months rather than digital months
  #ggplot(aes(x = sal_fb, y = temp_insitu), na.rm=TRUE) +
temp_sal <-  ggplot() +
  geom_point(data = d, aes(x = sal_fb, y = temp_fb, fill=month), na.rm=TRUE,
             size=size_point, show.legend=FALSE) + 
  labs(title="Temperature vs salinity, both in FerryBox, blue lines show density", x="Salinity", y="Temp. (°C)") +
  facet_grid(year ~ month) +
  stat_smooth(data = d, aes(x = sal_fb, y = temp_fb, color=month),
              method=lm, na.rm=TRUE, se=FALSE, show.legend=FALSE) +
  #scale_color_viridis(name="Month", discrete=TRUE, option="plasma") +
  geom_contour(data = grid, aes(x = s, y = t, z = z), 
               stat = "contour", binwidth=1, show.legend=FALSE, na.rm=TRUE) +
  scale_x_continuous(limits = c(30.4, 35.3), expand = c(0,0) ) +
  scale_y_continuous(limits = c(-2,8.5), expand = c(0,0) ) +
  coord_fixed() +
  #annotate("text", label="Fresh and cold", x=31.5, y=-1.5) +
  #annotate("text", label="Salty and warm", x=34.5, y=8.1) +
  Mytheme(size_labs = 10)
print(temp_sal)
ggsave(temp_sal,filename = paste0(path_fig, "temp_sal.png"), width = 20, height = 20 ,units="cm")
temp_sal_py <- ggplotly(temp_sal,dynamicTicks = TRUE)
filename.html <- paste0(path_fig, "temp_sal_py.html")
saveWidget(temp_sal_py, filename.html, selfcontained = T, libdir = "lib")
```

```{r temperature vs salinity regressions, echo=FALSE, message= FALSE, warning=FALSE, out.width="50%"}
# AT as a function of salinity for every year
j <- 0
reg <- data.frame()
for (i in 1:length(unique(d$year_month))) { #one loops through all years
  j <- j + 1
  ym <- unique(d$year_month)[i]
  d_reg <- filter(d, year_month==ym)
  if (length(!is.na(d_reg$temp_fb)) > 2) { #there is no temp_insitu in 2016-08, hence skip. Not needed if one uses temp_fb
  lms <- summary(lm(data=d_reg, sal_fb ~ temp_fb))
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <- pf(lms$fstatistic[1],lms$fstatistic[2],lms$fstatistic[3],lower.tail=FALSE)
  reg <- rbind(reg, as.numeric(c(lms$coefficients[2,1], lms$coefficients[2,2], 
                                 lms$coefficients[2,4], lms$coefficients[1,1], lms$coefficients[1,2],
                                 lms$coefficients[1,4], lms$fstatistic[1], lms$fstatistic[3], lms$r.squared, prob)))
  }
}
names(reg) <- c("Slope", "SE_Slope", "P Slope","Intercept","SE int.","P int.", "F", "df",
                "R2","P_value")
row.names(reg) <- as.character(unique(d$year_month))
#transform in latex table and save
tab_reg_temp_sal <- reg
tmp_reg_temp_sal <- reg %>% # to enable plotting below
  dplyr::mutate(date = as.Date(paste0(row.names(reg), "-01")),
                month = month(date, label=TRUE, abbr = TRUE))
tab_reg_temp_sal <- tab_reg_temp_sal[sort(rownames(tab_reg_temp_sal)),] # order chronologically

>>>>

d <- d_all %>%
  dplyr::filter(year < 2020) %>%
  dplyr::mutate(year_month = format(datetime, '%Y-%m')) %>%
  dplyr::mutate(month=month(datetime, label = TRUE, abbr = TRUE)) # abbreviated months rather than digital months
  #ggplot(aes(x = sal_fb, y = temp_insitu), na.rm=TRUE) +
temp_sal <-  ggplot() +
  geom_point(data = d, aes(x = sal_fb, y = temp_fb, fill=month), na.rm=TRUE,
             size=size_point, show.legend=FALSE) + 
  labs(title="Temperature vs salinity, both in FerryBox, blue lines show density", x="Salinity", y="Temp. (°C)") +
  facet_grid(year ~ month) +
  stat_smooth(data = d, aes(x = sal_fb, y = temp_fb, color=month),
              method=lm, na.rm=TRUE, se=FALSE, show.legend=FALSE) +
  #scale_color_viridis(name="Month", discrete=TRUE, option="plasma") +
  geom_contour(data = grid, aes(x = s, y = t, z = z), 
               stat = "contour", binwidth=1, show.legend=FALSE, na.rm=TRUE) +
  scale_x_continuous(limits = c(30.4, 35.3), expand = c(0,0) ) +
  scale_y_continuous(limits = c(-2,8.5), expand = c(0,0) ) +
  coord_fixed() +
  #annotate("text", label="Fresh and cold", x=31.5, y=-1.5) +
  #annotate("text", label="Salty and warm", x=34.5, y=8.1) +
  Mytheme(size_labs = 10)
print(temp_sal)
ggsave(temp_sal,filename = paste0(path_fig, "temp_sal.png"), width = 20, height = 20 ,units="cm")
temp_sal_py <- ggplotly(temp_sal,dynamicTicks = TRUE)
filename.html <- paste0(path_fig, "temp_sal_py.html")
saveWidget(temp_sal_py, filename.html, selfcontained = T, libdir = "lib")
```

```{r temperature vs salinity regressions, echo=FALSE, message= FALSE, warning=FALSE, out.width="50%"}
# AT as a function of salinity for every year
j <- 0
reg <- data.frame()
for (i in 1:length(unique(d$year_month))) { #one loops through all years
  j <- j + 1
  ym <- unique(d$year_month)[i]
  d_reg <- d %>%
    dplyr::select(year_month, sal_fb, temp_fb) %>%
    filter(year_month==ym, complete.cases(.))
  if (nrow(d_reg) > 2) { # There are months (2017-07) with no temp_fb nor sal_fb
  lms <- summary(lm(data=d_reg, sal_fb ~ temp_fb))
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <- pf(lms$fstatistic[1],lms$fstatistic[2],lms$fstatistic[3],lower.tail=FALSE)
  reg <- rbind(reg, as.numeric(c(lms$coefficients[2,1], lms$coefficients[2,2], 
                                 lms$coefficients[2,4], lms$coefficients[1,1], lms$coefficients[1,2],
                                 lms$coefficients[1,4], lms$fstatistic[1], lms$fstatistic[3], lms$r.squared, prob)))
  } else {
  reg <- rbind(reg, rep(NA, 9)) # if no sal and temp data fill with NA
  }
}
names(reg) <- c("Slope", "SE_Slope", "P Slope","Intercept","SE int.","P int.", "F", "df",
                "R2","P_value")
row.names(reg) <- as.character(unique(d$year_month))
#transform in latex table and save
tab_reg_temp_sal <- reg
tmp_reg_temp_sal <- reg %>% # to enable plotting below
  dplyr::mutate(date = as.Date(paste0(row.names(reg), "-01")),
                month = month(date, label=TRUE, abbr = TRUE))
tab_reg_temp_sal <- tab_reg_temp_sal[sort(rownames(tab_reg_temp_sal)),] # order chronologically
